DTOADQ

### 2017-10-09 documentation/_region_.tex
documentation & path tracing

\message{ !name(paper.tex)}\documentclass{article}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{color}
\usepackage[margin=1.15in]{geometry}

@@

  \maketitle
  \section{Introduction}
    I apologize for this being very lengthy and at times mathematically heavy,
    but there is no way of avoiding it.

  \section{SDFs}
  \begin{align*}
    S(P, R) &= P - R^2   & \text{SDF of a sphere Radius  }\\
    S(P, N) &= P \cdot N  & \text{SDF of a plane at Normal}\\
    S(P, H, R) &= <|P_{x, z}| - R, \: P_{y}> - H & \text{SDF of a torus with
                                                  Radius and Hole}\\
  \end{align*}
  \section{OpenCL}
  \section{Light Transport Introduction}
  \section{Bidirectional Path Tracing}
    All the materials in here, about path tracing, have come from Veach's PhD
    thesis. So, in a way, DTOADQ, at this stage, is just an implementation of
    BDPT as described in veach's thesis to the GPU, along with pulling from
    other resources such as Physically Based Render [PBR] 3rd ed. In order
    to understand the mathematics for BDPT, it must be worked up starting from
    light transport. So this section is about transforming the algorithms and
    mathematical models into something easily and efficiently implementable for
    the GPU, even if some of it is a rewrite of the mathematical model so that a
    programmer could understand it. In order to compute a normal of an SDF, you
    must perform a gradient approximation: the central distance onf the SDF at
    P. For example, the difference quotient:
      \[\frac{(f(x+h) - f(x))}{h}\]
    The most obvious way to calculate this is the six-point gradient; where for
    each axis you compute $f(x+h) - f(x-h)$, and normalize the result. This
    returns a vector pointing in the direction where the map's SDF changes the
    most: the normal. A four point gradient is possible; as long as each each
    point of the gradient is multiplied by $\epsilon$ before normalizing. The
    speedup from six to four is worthwhile, as computing each gradient is
    expensive as it involves a sdf map call. There are more utilities with
    gradients not yet currently provided in DTOADQ but are worth mentioning, a
    few are: an additional normal call using an alternative (higher-quality) SDF
    map can be used to compute a 'normal' bumpmap, and analyzing the gradient by
    hand allows a cheap normal computation of primitives such as a spheres.

@@

    And now
  \begin{align}
    F = \Sigma_s\Sigma_t \alpha^L_s c_{s, t} \alpha^E_t
  \end{align}
  \begin{align}
    \alpha^{L|E}_i =
    \begin{cases}
      \frac{L_e^0(y_0 \rightarrow y_1)}{p_A(y_1)} | \frac{W^0_e(z_0)}{p_A(z_0)},
      &i = 1,\\
      (\frac{f_s(y_{i+1} \rightarrow y_i \rightarrow y_{i-1})}{p_{\sigma}(y_i
      \rightarrow y_{i-1})} | \frac{f_s(z_{i-1} \rightarrow z_{i-2} \rightarrow
      z_{i-3}}{p_\sigma(z_{i-2} \rightarrow z_{i-1})}) \alpha^{L|E}_{i-1},
      &i > 1
    \end{cases}
  \end{align}
    There is no case for $i = 0$ as either, the path doesn't exist, or if one
    does exist, $y_0$ and $z_0$ do not contribute to the image. Veach (and many other
    sources), leave this in, but it's unnecessary and can be confusing. However,
    there may be cases where $s = 0$ ($t = 0$ is not possible without a
    physical camera lens), which is equivalent to just forward path tracing.
    Well, in either case where both paths exist or only the camera path exists,
    the connection strategy below describes how to connect the edges of the two
    paths
  \begin{align}
    c_{s, t} =
    \begin{cases}
      L_e(z_{-2} \rightarrow z_{-1}), &s = 0, t > 0,\\
      f_s(y_{-2} \rightarrow y_{-1} \rightarrow z_{-1}) G(y_{-1} \leftrightarrow
      z_{-1}) f_s(y_{-1} \rightarrow z_{-1} \rightarrow z_{-2}), &s = s, t > 0
    \end{cases}
  \end{align}
    One immediately obvious optimization that could be made with this model is
    in regards to the visibility check in the geometric term, if you were to
    expand $I_j$, and take the V term from outside the G function
    [$G(v \rightarrow
    \dot{v}) V(v \rightarrow \dot{v})$],
    it would be made obvious that all the visibility checks can be
    cancelled out for the entire equation, except for the connection term $c$.
    The expansion is not shown as it is very lengthy, and unnecessary. Paths
    $\bar{y}_{0} \ldots \bar{y}_s$ and $\bar{z}_0 \ldots \bar{z}_t$ being
    visible is intuitively obvious as we are already generating these paths.
    Specifically, the definition of $V$ is
  \begin{align}
    V(v \rightarrow \dot{v}) = ||v - \dot{v}|| \leq
    m(v,\:\overrightarrow{v\:\dot{v}})
  \end{align}
    where $m$ is a march through the SDF map. Since the paths are being
    generated using march. A visibility check is only necessary for the
    connection edge
    \begin{align*}
      ||v - \dot{v}|| = m(v, \overrightarrow{v, \dot{v}})\\
      \:\:\:\text{for all paths but } y_{-1} \rightarrow z_{-1} 
    \end{align*}
\\
    Moving on, for $\alpha^{L|E}_i$, these are equivalent for both $L$ and $E$,
    the problem is that the $\bar{y}$ path is generated in perspective of the
    light source, so we have to call the BSDF in perspective of the camera; the
    path is flipped. Handling $\bar{y}$ is significantly easier than $\bar{z}$,
    as $\bar{y}$ is precomputed. For $\bar{z}$, the next vertex $z_{+1}$
    ($\omega_o$ in the perspective of SA), is unknown. For example $\alpha^L_1$
    in OpenCL looks like:

@@

    For $c_{s, t}$, there is no cases for $s = s, t = 0$ as
    the lens does not have a physical existence in the SDF map. $p_A$ is
    the PDF w.r.t. area, for the light path; we only need to take the PDF of the
    solid angle of a sphere
  \begin{align}
    p_A^L(y_0 \rightarrow y_1) = p_{\sigma}^L(y_0 \rightarrow y_1) G(v
    \leftrightarrow \dot{v})\\
    p_{\sigma}^L(y_0 \rightarrow y_1) = \frac{1.0f}{\tau * (1.0f -
    \sqrt{\frac{R^2}{|y_0 - y_1|^2}})}
  \end{align}
    where $R$ is the radius of the sphere, note $y_0$ is a point on the surface
    of the sphere, and not its origin.
    \\$p_A^E(x_0 \rightarrow x_1)$ doesn't exist, as there is no physical lens
    on the scene. [TODO - this might cahnge later]. One of the benefits of this,
    is that there are no special cases for pathes where $t = 0$, which is
    impossible to handle on the GPU as the path could contribute to any
    arbitrary pixel [directly hitting the lens].
  \begin{align}
    P_A^E = 1.0f
  \end{align}
    The $L^0_e$ is contribution from the light source. As of right now, DTOADQ
    only supports area lights, so $L^0_e$ is set a constant for the emitter's
    spectrum value.
    The $W_e^0$ is contribution from the eye source, that is, the lens. Since
    the lens does not exist in the scene, as of now [TODO - this might change
    later], there is no $W_e^0$. While this should be implemented in the future,
    a hidden benefit of the current model is that there are no special cases for
    paths where $t = 0$, which is impossible to handle on the GPU as any such
    path could contribute to any arbitrary pixel [the light path is, after all,
    directly hitting the lens]. So this has now become:
  \begin{align}
    \alpha_i^E = \Sigma_{i\leq 0} 
  \end{align}
  The amount that each path contributes to F differ greatly, so naively
  weighing each contribution uniformly by something like 1.0f, or even, by
  inverse path length, will producy very noisy images, and would not be
  worth the additional computation time that BDPT requires. In order to get good
  samples from the contributions, they must be combined in an optimal manner
  using Multiple Importance Sampling: each contribution is weighted by the PDF
  of the entire path itself:
  \begin{align*}
    F &\equiv \Sigma_s\Sigma_tW_{s, t}C_{s, t}\\
    C_{s, t} &= \alpha^L_s c_{s, t} \alpha^E_t
  \end{align*}
    where $W_{s, t}$ is the MIS weight of path $\bar{y} \leftarrow \bar{z}$, and
    $C_{s, t}$ is the unweighted contribution of the same path. For clarity,
    $C^U_{s, t}$ as described below is the unweighted contribution of the
    algorithm:
  \begin{align}
    C^u_{s, t} = (\alpha^L_s c_{s, t} \alpha^E_t) \equiv
    \frac{f_j(\bar{v})}{p(\bar{v})}
  \end{align}
  thus
  \begin{align}
    F = \Sigma_s\Sigma_tW_{s, t}C^u_{s, t}
  \end{align}
    The formula Veach gives for Multiple Importance Sampling on BDPT is as so:
  \begin{align}
    MIS_{st}(\bar{x}) = \frac{P_s(\bar{x})}{\Sigma_i{P_i(\bar{x})}} 
  \end{align}
    Conceptually, the numerator $P_s(\bar{x})$ is the path density that was
    generated, while the denominator is the path density of other connection
    strategies that could could have, in theory,$^{[1]}$ created the path. The
    straight forward implementation of this has a lot of problems gone
    unresolved in Veach's thesis; PBR 3rd Ed, chapter 16 pg 1014-1016 addresses
    and describe these problems. Primarily, $P_i(\bar{x})$ will overflow
    CLFloat (due to the distance in the Geometric term), and the straight
    forward implementation is very long, hard to debug, and has a poor time
    complexity. But the important part is the end result of their solution:

@@

@@

\message{ !name(paper.tex) !offset(-368) }

>>>>

\message{ !name(paper.tex)}\documentclass{article}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{color}
\usepackage{graphicx}
\graphicspath{ {../screenshots/} }
\usepackage[margin=1.15in]{geometry}

@@

  \maketitle
  \section{Introduction}

  \section{SDFs}
  \begin{align*}
    S(P, R) &= P - R^2   & \text{SDF of a sphere Radius  }\\
    S(P, N) &= P \cdot N  & \text{SDF of a plane at Normal}\\
    S(P, H, R) &= <|P_{x, z}| - R, \: P_{y}> - H & \text{SDF of a torus with
                                                  Radius and Hole}\\
  \end{align*}
  \section{OpenCL}
  \section{Light Transport Introduction}
  \section{Bidirectional Path Tracing}
    All the materials in here, about path tracing, have come from Veach's PhD
    thesis. So, in a way, DTOADQ, at this stage, is just an implementation of
    BDPT as described in veach's thesis to the GPU, along with pulling from
    other resources such as Physically Based Render [PBR] 3rd ed. In order
    to understand the mathematics for BDPT, it must be worked up starting from
    light transport. So this section is about transforming the algorithms and
    mathematical models into something easily and efficiently implementable for
    the GPU, even if some of it is a rewrite of the mathematical model so that a
    programmer could understand it. In order to compute a normal of an SDF, you
    must perform a gradient approximation: the central distance onf the SDF at
    P. For example, the difference quotient:
      \[\frac{(f(x+h) - f(x))}{h}\]

    Calculating probability distribution functions is a very vital component in
    Monte Carlo Sampling. The PDF states the probability of a random variable to
    take the form of a given input. Its usefulness comes when contributions from
    many Monte Carlo samples have a different weight. That is, one sample may
    have a higher probability distribution, and thus a higher contribution to
    the path. That's the strategy behind Monte Carlo;
    \begin{align}
      \int f(x) \approx \frac{1}{N}\Sigma_0^N\frac{f(x)}{p(x)}
    \end{align}
    The PDF applies to everything from path contributions, solid angles of a
    BRDF, uniformly sampling a sphere, etc. I had a difficult time understanding
    these, and more importantly, how to calculate them for a given formula. The
    biggest conceptual barrier is that, for example, given a solid angle in a
    hemisphere $\Omega$, the probability for a random variable X to equal
    $\omega$ is 0. But this isn't what's happening here, this is all about the
    probability that X could potentially equal $\omega$ when compared to other values
    in $\Omega$. It's all about comparing the probabilities of two potential
    $\omega$s with each other.

    I still have a hard time conceptualizing these things in my head, but the
    good thing is, is that calculating these are relatively easy. There's a lot
    that could be talked about, but the important thing is that I've found a
    reliable three-step process. The first is to
    find the cumulative distribution function $CDF = Pr(X \leq x)$. Where $Pr$
    is the probability, $X$ is a uniformly random variable, and $x$ is a known
    constant. In other words, this gives the probability that, for a formula
    $P$, where $X \in \Omega$, and $\Omega$ is the set of all possible values
    for $CDF$,
    that $X \leq x$. The second step is to take the derivative of this, so for
    PDF $p$
    \begin{align}
      p(x) = \frac{dCDF(x)}{dx}
    \end{align}
    And the final step is to verify that the PDF is correct by integrating it
    over $\Omega$, this final value should equal 1. This makes sense intuitively
    -- it is a probability after all, so the sum of the probabilities of every
    possible input must be 1.
    For an example, how I calculated the PDF of an area light source
      $L^0_e(v) = L_o(P, \omega_o)$ 
    is quite simple. Firstly,the area light is defined as a sphere, and thus our
    set of all possible values is correlated to surface area:
    $\Omega = [0,\:2.0f*\tau*R^2]$.
    Thus for a given constant $x$, and a uniformly random $X$,
    \begin{align}
      CDF(x) = Pr(X \leq x) = \frac{x}{2.0f*\tau*R^2}\\
      p(x) = \frac{dCDF(x)}{dx} = \frac{1.0f}{2.0f*\tau*R^2}\\
      \int_0^{\Omega}p(x)dx = 1.0f
    \end{align}
    In code
    \begin{lstlisting}
      float Light_PDF ( float3 P, Emitter* m ) {
        float rad_sqr = m->radius*m->radius;
        return 1.0f/(2.0f*TAU*rad_sqr);
      }
    \end{lstlisting}
    The more interesting aspect, in this example, is actually generating the
    uniformly distributed samples of a sphere. The naive method gives incorrect
    distributions.TODO {page 774}

    The most obvious way to calculate this is the six-point gradient; where for
    each axis you compute $f(x+h) - f(x-h)$, and normalize the result. This
    returns a vector pointing in the direction where the map's SDF changes the
    most: the normal. A four point gradient is possible; as long as each each
    point of the gradient is multiplied by $\epsilon$ before normalizing. The
    speedup from six to four is worthwhile, as computing each gradient is
    expensive as it involves a sdf map call. There are more utilities with
    gradients not yet currently provided in DTOADQ but are worth mentioning, a
    few are: an additional normal call using an alternative (higher-quality) SDF
    map can be used to compute a 'normal' bumpmap, and analyzing the gradient by
    hand allows a cheap normal computation of primitives such as a spheres.

@@

    And now
  \begin{align}
    F = \Sigma_s\Sigma_t \alpha^L_s c_{s, t} \alpha^E_t
  \end{align}
  \begin{align}
    \alpha^{L|E}_i =
    \begin{cases}
      \frac{L_e^0(y_0 \rightarrow y_1)}{p_A(y_1)} | \frac{W^0_e(z_0)}{p_A(z_0)},
      &i = 1,\\
      (\frac{f_s(y_{i+1} \rightarrow y_i \rightarrow y_{i-1})}{p_{\sigma}(y_i
      \rightarrow y_{i-1})} | \frac{f_s(z_{i-1} \rightarrow z_{i-2} \rightarrow
      z_{i-3}}{p_\sigma(z_{i-2} \rightarrow z_{i-1})}) \alpha^{L|E}_{i-1},
      &i > 1
    \end{cases}
  \end{align}
    There is no case for $i = 0$ as either, the path doesn't exist, or if one
    does exist, $y_0$ and $z_0$ do not contribute to the image. Veach (and many other
    sources), leave this in, but it's unnecessarily confusing. A special case
    needs to be handled, unfortunately, for $\alpha^E_0$ as it is being
    generated on the fly while $\bar{z}$ is being constructed. The two options
    is to unroll the first iteration of the construction loop, or just allow the
    special case to exist. The latter is the better choice, as all kernels will
    enter and exit the special case at the same time, and the GPU might just
    unroll the loop anyways.
    There may be cases where $s = 0$ ($t = 0$ is not possible without a
    physical camera lens), which is equivalent to just forward path tracing.
    There is also to consider, what happens on $s = 1$ and $t = 1$, %TODO
    Well, in either case where both paths exist or only the camera path exists,
    the connection strategy below describes how to connect the edges of the two
    paths
  \begin{align}
    c_{s, t} =
    \begin{cases}
      L_e(z_{-2} \rightarrow z_{-1}), &s = 0, t > 0,\\
      f_s(y_{-2} \rightarrow y_{-1} \rightarrow z_{-1}) G(y_{-1} \leftrightarrow
      z_{-1}) f_s(y_{-1} \rightarrow z_{-1} \rightarrow z_{-2}), &s > 0, t > 0
    \end{cases}
  \end{align}
    One question to be made is if the strategy $s = 0, t > 0$ is worthwhile to
    compute; the most noticeable effect are for caustics, light sources in
    direct view of an emitter should be emitted from animations in this current
    state of DTOADQ. Another idea to consider is that under real physics
    circumstances, the photons are still allowed to bounce after hitting a
    source of light, but the contribution from these cases are negligible, and
    so most likely not worthwhile to implement. Anyways, the $s = 0$
    strategy is a worthwhile  calculation under these circumstances

    One immediately obvious optimization that could be made with this model is
    in regards to the visibility check in the geometric term, if you were to
    expand $I_j$, and take the V term from outside the G function
    [$G(v \rightarrow
    \dot{v}) V(v \rightarrow \dot{v})$],
    it would be made obvious that all the visibility checks can be
    cancelled out for the entire equation, except for the connection term $c$.
    The expansion is not shown as it is very lengthy, and unnecessary. Paths
    $\bar{y}_{0} \ldots \bar{y}_s$ and $\bar{z}_0 \ldots \bar{z}_t$ being
    visible is intuitively obvious as the paths have been generated using the
    same technique used to check for visibility; raymarching. Specifically, the
    definition of $V$ is
  \begin{align}
    V(v \rightarrow \dot{v}) = ||v - \dot{v}|| \leq
    m(v,\:\overrightarrow{v\:\dot{v}})
  \end{align}
    where $m$ is a march through the SDF map. A visibility check is only
    necessary for the connection edge
    \begin{align*}
      ||v - \dot{v}|| = m(v, \overrightarrow{v, \dot{v}})\\
      \:\:\:\text{for all paths but } y_{-1} \rightarrow z_{-1} 
    \end{align*}
\\
    Moving on, for $\alpha^{L|E}_i$, these are equivalent for both $L$ and $E$,
    the problem is that the $\bar{y}$ path is generated in perspective of the
    light source, so we have to call the BSDF in perspective of the camera; the
    path is flipped. Handling $\bar{y}$ is significantly easier than $\bar{z}$,
    as $\bar{y}$ is precomputed. For $\bar{z}$, the next vertex $z_{+1}$
    ($\omega_o$ in the perspective of SA), is unknown. For example $\alpha^L_1$
    in OpenCL looks like:

@@

    For $c_{s, t}$, there is no cases for $s = s, t = 0$ as
    the lens does not have a physical existence in the SDF map. $p_A$ is
    the PDF w.r.t. area, for the light path; we only need to take the PDF of the
    solid angle of a sphere
  \begin{align}
    p_A^L(y_0 \rightarrow y_1) = p_{\sigma}^L(y_0 \rightarrow y_1) G(v
    \leftrightarrow \dot{v})\\
    p_{\sigma}^L(y_0 \rightarrow y_1) = \frac{1.0f}{\tau * (1.0f -
    \sqrt{\frac{R^2}{|y_0 - y_1|^2}})}
  \end{align}
    where $R$ is the radius of the sphere, note $y_0$ is a point on the surface
    of the sphere, and not its origin.
    \\$p_A^E(x_0 \rightarrow x_1)$ doesn't exist, as there is no physical lens
    on the scene. [TODO - this might cahnge later]. One of the benefits of this,
    is that there are no special cases for pathes where $t = 0$, which is
    impossible to handle on the GPU as the path could contribute to any
    arbitrary pixel [directly hitting the lens].
  \begin{align}
    P_A^E = 1.0f
  \end{align}
    The $L^0_e$ is contribution from the light source. As of right now, DTOADQ
    only supports area lights, so $L^0_e$ is set a constant for the emitter's
    spectrum value.
    The $W_e^0$ is contribution from the eye source, that is, the lens. Since
    the lens does not exist in the scene, as of now [TODO - this might change
    later], there is no $W_e^0$. While this should be implemented in the future,
    a hidden benefit of the current model is that there are no special cases for
    paths where $t = 0$, which is impossible to handle on the GPU as any such
    path could contribute to any arbitrary pixel [the light path is, after all,
    directly hitting the lens]. So this has now become:
  \begin{align}
    \alpha_i^E = \Sigma_{i\leq 0} 
  \end{align}
  The amount that each path contributes to F differ greatly, so naively
  weighing each contribution uniformly by something like 1.0f, or even, by
  inverse path length, will producy very noisy images, and would not be
  worth the additional computation time that BDPT requires, as shown below.
  \begin{align}
 \includegraphics{bdpt_unweighted}
  \end{align}
  In order to get good
  samples from the contributions, they must be combined in an optimal manner
  using Multiple Importance Sampling: each contribution is weighted by the PDF
  of the entire path itself:
  \begin{align*}
    F &\equiv \Sigma_s\Sigma_tW_{s, t}C_{s, t}\\
    C_{s, t} &= \alpha^L_s c_{s, t} \alpha^E_t
  \end{align*}
    where $W_{s, t}$ is the MIS weight of path $\bar{y} \leftarrow \bar{z}$, and
    $C_{s, t}$ is the unweighted contribution of the same path. For clarity,
    $C^U_{s, t}$ as described below is the unweighted contribution of the
    algorithm:
  \begin{align}
    C^u_{s, t} = (\alpha^L_s c_{s, t} \alpha^E_t) \equiv
    \frac{f_j(\bar{v})}{p(\bar{v})}
  \end{align}
  thus
  \begin{align}
    F = \Sigma_s\Sigma_tW_{s, t}C^u_{s, t}
  \end{align}
    The formula for Multiple Importance Sampling on BDPT is as so:
  \begin{align}
    MIS_{st}(\bar{x}) = \frac{P_s(\bar{x})}{\Sigma_i{P_i(\bar{x})}} 
  \end{align}
    Conceptually, the numerator $P_s(\bar{x})$ is the path density that was
    generated, while the denominator is the path density of other connection
    strategies that could could have, in theory,$^{[1]}$ created the path. The
    straight forward implementation of this has a lot of problems gone
    unresolved in Veach's thesis; PBR 3rd Ed, chapter 16 pg 1014-1016 addresses
    and describe these problems. Primarily, $P_i(\bar{x})$ will overflow
    CLFloat (due to the distance in the Geometric term), and the straight
    forward implementation is very long, hard to debug, and has a poor time
    complexity. But the important part is the end result of their solution:

@@

  \section{Future Improvements}

  Every emitter is chosen uniformly, however a better system would weight the
  probability of each emitter by
  \[
    p_l = \frac{\phi_lr_l}{\Sigma_i\phi_ir_l}
  \]
  Thus emitters that have a higher contribution to the scene, that is, a larger
  emittance and radius, will have a higher probability to be selected.

@@

\message{ !name(paper.tex) !offset(-461) }
